Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-04-17 doc/invt.tex
small edits invt.tex

We develop the theory
in the following order.
%
In Sec. \ref{sec:background},
we first review the basics of FDS
and some known aspects of the optimal schedule.
%
%This also helps fixing the notations.
%
Then, in Sec. \ref{sec:single-bin},
we illustrate the method of
computing the optimal schedule
in the simple case of the
single-bin updating scheme
used by the WL algorithm,
by proving the optimality
of the known inverse-time formula.
%
Next, we extend the method to the general case
of a multiple-bin updating scheme
that encompasses metadynamics
in Secs. \ref{sec:multiple-bin}
to \ref{sec:band-matrix}.
%
Finally, we compare different updating schemes
and show that the single-bin scheme
and a class of generalized bandpass schemes
are optimal for convergence
in the long time or extensive sampling limit
in Sec. \ref{sec:cmpschemes}.

@@

For a large system at temperatures of interest,
the distribution, $p^*_i$, often tends to
be localized,
%
and to explore the global shape
of the PMF, $-\ln p^*_i$,
it is often advantageous to perform
biased sampling that targets
a wider and smoother distribution, $p_i$.
%
%Here, we refer to simulations that target
%a flat or nearly flat distribution
%as entropic or multicanonical sampling.

To do so, we introduce a bias potential, $V_i$,
which changes the equilibrium distribution to
%
\begin{equation}
  \pi_i \propto p^*_i \, e^{-V_i}.
  \label{eq:pi_p_V1}
\end{equation}
%
Upon normalization,
$\sum_{i = 1}^n \pi_i = 1$,
we get
%
\begin{equation}
  \pi_i
  =
  \frac{                p^*_i \, e^{-V_i} }
       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
  .
  \label{eq:pi_p_V}
\end{equation}
%
%In particular,
%to make $\pi_i$ flat,
%the bias potential $V_i$
%must coincide with $\ln p^*_i$,
%up to an additive constant.
%%
%In practice, $V_i$ may be non-optimal or contain error,
%and thus $\pi_i$ need not
%coincide with the intended distribution $p_i$.

We wish the biased equilibrium distribution, $\pi_i$,
to converge to the target distribution, $p_i$,
which is usually chosen to be flat,
i.e. $p_i$ is constant for all $i$'s,
or nearly flat\cite{
dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}.
%
We will broadly call such a biased simulation
a flat-distribution-sampling (FDS) simulation.
%
To achieve this distribution, the bias potential must satisfy
%
\begin{equation}
  V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
  ,
  \label{eq:Vi_target}
\end{equation}
%
according to Eq. \eqref{eq:pi_p_V1}.
%which allows us to deduce the PMF, $-\ln p^*_i$,
%directly from the bias potential.
%
This relation thus sets a practical aim for the convergence of FDS
methods---the bias potential must be adjusted to make the
observed distribution, or the normalized histogram,
match the intended one.
%
Once this aim is achieved,
the PMF, $-\ln p^*_i$, can be naturally
deduced from the bias potential. % $V_i$.
\note{
  For example, in the context of simulated tempering (Zhang and Ma, 2007),
  $p^*_i$ is replaced by $\ln Z_i$,
  $p_i$ is the weight $\zeta_i$,
  and the bias potential, $V_i$, is $\ln (\tilde Z_i / \zeta_i)$.
  Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
  up to a constant shift.
}

Roughly speaking, there are two types of FDS methods.
%
The first type,
referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
uses a fixed bias potential, $V_i$,
(estimated a priori)
during simulation.
%
Examples include the original
entropic or multicanonical sampling\cite{
berg1992, *lee1993}
and simulated tempering\cite{
marinari1992, *lyubartsev1992}
with $i$ being the index of energy
and that of temperature, respectively.
%
The estimated bias potentials at bin $i$, $V_i$,
are often improved
with the normalized histogram of the sampling, $H_i$, accumulated
during simulation.
%
Using $H_i$ as an estimate of $\pi_i$ in
Eq. \eqref{eq:pi_p_V1}
allows us to correct $V_i$ as
%
\begin{equation}
  V^\mathrm{corr}_i
  =
  V_i
  +
  \ln \frac{ H_i }
           { p_i }.
  \label{eq:vcorr_equil}
\end{equation}
%
\note{This follows from
  $$
  \begin{aligned}
    H_i \approx \pi_i
    &\propto p^*_i \, e^{-V_i},
    \\
    p_i
    &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
  \end{aligned}
  $$
}
However, for an unknown system,
it is often difficult
to guess a sufficiently accurate
$V_i$ with which to start.
%
Thus,
it can take many iterative rounds of simulations
to achieve the desired flat distribution.
%
%Another difficulty of this approach is that
The effectiveness of this approach
depends on the precision of the histogram, $H_i$,
which in turn depends on the simulation length
that must be decided in advance.
%
Thus, the rigidity of this framework disallows
continuous, rapid improvement of the bias potential,
and subsequently the PMF,
as simulation lengthens.

The second type of methods,
which we will refer to as the adaptive FDS methods,
address the above shortcomings
by continuously and incrementally updating the bias potential.
%For this reason, the second type of methods,
%referred to as the adaptive FDS methods,
%were invented.
%
Examples include
the WL algorithm and metadynamics.
%
%These methods incrementally update the bias potential
%and can approximately sample the target distribution.
%
However,
the on-the-fly updates break the microscopic reversibility
of the underlying sampling mechanism.
Further, failure to properly reduce the magnitude of
the histogram-based increments
would lead to roughness or errors in the final estimate of the PMF.
%
%Thus, for a conservative practitioner,
%the E-FDS methods appear more appealing
%once a sufficiently accurate bias potential is available.
%
To minimize this effect,
many adaptive methods adopt a schedule to
reduce the updating magnitude
over the course of simulation\cite{
marsili2006,
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007,
barducci2008}.
%
%Thus, in late stages of a long simulation,
%the updating becomes negligible and
%the adaptive simulation tends to
%the corresponding equilibrium one.
%
%if the bias potential were converged to the exact value,
%this adaptive FDS tends to the ideal equilibrium counterpart,
%particular equilibrium simulation can be thought as
%the optimal asymptotic limit of the adaptive counterpart,
%
For example, one can show that
if the magnitude is controlled by
the optimal inverse-time prescription\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
liang2007,
morozov2007, zhou2008,
komura2012, *caparica2012, *caparica2014},
the WL algorithm is equivalent to the ideal equilibrium counterpart
in improving the precision of the bias potential,
thus allowing for maximum efficiency
(cf. Appendix \ref{sec:equilerr}).
%the bias potential in the WL algorithm
%will evolve similar to Eq. \eqref{eq:vcorr_equil}
%with the precision of the histogram as simulation lengthens.

@@

We will mainly focus on the adaptive methods below.
%
In the WL algorithm\cite{wang2001, *wang2001pre},
the bias potential, $V_i(t)$, is updated
at each sampling step $t$ as
%
\begin{equation}
  V_i(t+1)
  =
  V_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { p_{i(t)} }
  ,
\label{eq:wl_update}
\end{equation}
%
where $i(t)$ is the bin at step $t$,
$\delta_{i, \, i(t)}$ is the Kronecker delta function,
which is $1$ when $i = i(t)$ or $0$ otherwise,
and $\alpha(t)$ is the updating magnitude.
%
We refer to this scheme as the \emph{single-bin scheme},
for it applies only to the bias potential
at the current bin, $i(t)$.
%
In a more general \emph{multiple-bin scheme}, however,
several neighboring bins are updated:
%
\begin{equation}
  V_i(t+1)
  =
  V_i(t)
  +
  w_{i, \, i(t)}
  \frac{ \alpha(t) }
       { p_{i(t)} },
  \label{eq:mbin_update}
\end{equation}
%
where the elements, $w_{ij}$'s, are a set of coefficients satisfying
the restrictions to be discussed
in Sec. \ref{sec:updating-matrix}.
%
The Gaussian updating scheme
used in metadynamics
is a multiple-bin scheme.

@@

%For the single-bin scheme
%used in WL algorithm,
Later,
it was found that
a better way
of controlling the schedule, $\alpha(t)$,
in late stages
is to follow the formula
%
\begin{equation}
  \alpha(t) = \frac{1}{t},
  \label{eq:alpha_invt}
\end{equation}
%
where $t$,
referred to as the ``time'' below,
is the number of steps
from the beginning of the simulation.
%
This simple yet effective formula has attracted
several studies\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
morozov2007, zhou2008,
komura2012, *caparica2012, *caparica2014}.
%
An interesting explanation\cite{
  marsili2006, barducci2008}
comes from the interpretation of
the bias potential %under the WL updating scheme, Eq. \eqref{eq:wl_update},
as a runtime average computed from all data points collected so far,
so that the addition of a new data point only
affects the average by a weight of $1/t$,
which is the inverse of the current sample size, $t$
(cf. Appendix \ref{sec:equilerr}).
%
Here we will re-derive this formula
and suggest a slight modification
whereby the origin of $t$ is shifted
according to the initial error.
%
Interestingly, we may reach the same result
from optimally planning the simulation lengths in the stages
and the reduction factor, as shown in Appendix \ref{sec:wltoinvt}.

As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
several similar ways of
reducing the updating magnitude
were developed for umbrella sampling and metadynamics,
particularly, the self-healing umbrella sampling
and well-tempered metadynamics\cite{
  marsili2006, barducci2008, dickson2011, dama2014}.
%
Our aim here is to find the optimal schedule, $\alpha(t)$,
for a general updating scheme.

@@

First, we need to define the error in the bias potential.
%
This is done by deducting two contributions from the bias potential.
%
The first contribution comes from the target value,
Eq. \eqref{eq:Vi_target}.
%
For this purpose,
we introduce a shifted bias potential
%
\begin{equation}
  v_i(t)
  \equiv
  V_i(t)
  -
  \ln \frac { p^*_i }
            { p_i }
  ,
  \label{eq:v_def}
\end{equation}
%
which should tend to a constant of $i$
upon convergence.
Note that the updates to $v_i$'s are
equivalent to the updates to $V_i$'s,
since the shift
remains a constant during the course of simulation,
%%
%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
%becomes
%%
%\begin{equation}
%  \pi_i
%  =
%  \frac{                p_i \, e^{-v_i} }
%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
%  \propto
%  p_i \, e^{-v_i}.
%  \label{eq:pi_p_v}
%\end{equation}
%%
%According to Eq. \eqref{eq:pi_p_V1},

@@

This is a functional of the curve $q(t)$.
%
We now fix the endpoint value, $q(T)$,
and vary the curve, $q(t)$ ($0 < t < T$).
%
This is equivalent to varying $u$
with fixed endpoints,
$u\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$u\bigl( q(T) \bigr) = 1$.
%
Note also that the first term on the right-hand side
of Eq. \eqref{eq:x2t_average} is fixed during
the process,
while the second term is the action of a free particle,
which must move at a constant velocity, $c$:
%
%We will further demand that
%%
%\begin{equation}
%  \lim_{t \to \infty} q(t) \to \infty.
%  \label{eq:qt_limit}
%\end{equation}
%%
%Then, for a large $t$,
%the second term on the right-hand side
%of Eq. \eqref{eq:xt_solution} can be neglected, and
%%
%\begin{align}
%\left\langle x^2(t) \right\rangle
%%&=
%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%%\notag
%%\\
%&=
%\int_0^t \int_0^t
%  \dot u\bigl( q(t') \bigr) \,
%  \dot u\bigl( q(t'') \bigr) \,
%  \kappa(t' - t'') \, dt'' \, dt'.
%\label{eq:x2t_average_asym}
%\end{align}
%
%By the Euler-Lagrange equation, we get
%$$
%\begin{aligned}
%0
%&=
%\frac{d}{dt} \int_0^T
%  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
%\\
%&= \int_0^T
%  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
%\end{aligned}
%$$
%where we have dropped the boundary terms
%by assuming Eq. \eqref{eq:kappa_limit}
%and a large $T$.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
%In order for this to hold for any $t$,
%we must have
%%
%$
%\ddot u\bigl( q(t') \bigr) = 0
%,
%$
%or
\begin{equation}
  \dot u\bigl( q(t') \bigr) = c
  .
\label{eq:dudt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:u_def}, we have
$$
e^{ q(t') - q(T) }
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
%
Taking the logarithm,
and differentiating both sides
with respect to $t'$,
we get Eq. \eqref{eq:alpha_invt1}.
%
From the boundary condition at $t' = T$,
we get $c = 1/(T + t_0)$,
and from the case of $t' = 0$,
we get $q(T) = \ln(1 + T/t_0)$.
%
Since the above optimization is performed
under a fixed $q(T)$, below
one will choose a proper value of $q(T)$,
or, equivalently, of $t_0$, to complete the optimization.

>>>>

We develop the theory
in the following order.
%
In Sec. \ref{sec:background},
we first review the basics of FDS
and some known aspects of the optimal schedule.
%
Then, in Sec. \ref{sec:single-bin},
we illustrate the method of
computing the optimal schedule
in the simple case of the
single-bin updating scheme
used by the WL algorithm,
by proving the optimality
of the known inverse-time formula.
%
The above two introductory sections help fix our notations,
and an expert reader can readily skip them.
%
Next, we extend the method to the general case
of a multiple-bin updating scheme
that encompasses metadynamics
in Secs. \ref{sec:multiple-bin}
to \ref{sec:band-matrix}.
%
Finally, we compare different updating schemes
and show that the single-bin scheme
and a class of generalized bandpass schemes
are optimal for convergence
in the long time or extensive sampling limit
in Sec. \ref{sec:cmpschemes}.

@@

%For a large system at temperatures of interest,
%the distribution, $p^*_i$, often tends to
%be localized,
%%
%and to explore the global shape
%of the PMF, $-\ln p^*_i$,
%it is often advantageous to perform
%biased sampling that targets
%a wider and smoother distribution, $p_i$.
%%
%%Here, we refer to simulations that target
%%a flat or nearly flat distribution
%%as entropic or multicanonical sampling.
%
%To do so, we introduce a bias potential, $V_i$,
If the distribution, $p^*_i$, is multimodal,
it is often advantageous to perform
biased sampling that targets
a smoother distribution, $p_i$,
by introducing a bias potential, $V_i$,
such that the biased equilibrium distribution is
%
\begin{equation}
  \pi_i
  %=
  %C_i \, p^*_i \, e^{-V_i}
  =
  \frac{                p^*_i \, e^{-V_i} }
       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
  \propto
  p^*_i \, e^{-V_i}
  .
  \label{eq:pi_p_V}
\end{equation}
%
%where the proportionality constant, $C_i$,
%is found from the normalization condition,
%$\sum_{i = 1}^n \pi_i = 1$,
%in the second step.
%
%In particular,
%to make $\pi_i$ flat,
%the bias potential $V_i$
%must coincide with $\ln p^*_i$,
%up to an additive constant.
%%
%In practice, $V_i$ may be non-optimal or contain error,
%and thus $\pi_i$ need not
%coincide with the intended distribution $p_i$.

We will broadly call a biased simulation
targeting a flat (i.e. $p_i \equiv 1/n$)
or nearly flat\cite{
dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}
$p_i$
a flat-distribution-sampling (FDS) simulation.
%
For the biased distribution, $\pi_i$,
to converge to $p_i$,
the bias potential must satisfy
%
\begin{equation}
  V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
  .
  \label{eq:Vi_target}
\end{equation}
%
%according to Eq. \eqref{eq:pi_p_V}.
%which allows us to deduce the PMF, $-\ln p^*_i$,
%directly from the bias potential.
%
In practice,
the bias potential is adjusted to make the
observed distribution, or the normalized histogram,
match the intended one,
and once this is done,
the PMF, $-\ln p^*_i$, can be naturally
deduced from the bias potential. % $V_i$.
%
%This relation thus sets a practical aim for the convergence of FDS
%methods---the bias potential must be adjusted to make the
%observed distribution, or the normalized histogram,
%match the intended one.
%%
%Once this aim is achieved,
%the PMF, $-\ln p^*_i$, can be naturally
%deduced from the bias potential. % $V_i$.
\note{
  For example, in the context of simulated tempering (Zhang and Ma, 2007),
  $p^*_i$ is replaced by $\ln Z_i$,
  $p_i$ is the weight $\zeta_i$,
  and the bias potential, $V_i$, is $\ln (\tilde Z_i / \zeta_i)$.
  Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
  up to a constant shift.
}

Roughly speaking, there are two types of FDS methods.
%
The first type,
referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
uses a fixed bias potential, $V_i$,
(estimated a priori)
during simulation.
%
Examples include the original
entropic or multicanonical sampling\cite{
berg1992, *lee1993}
and simulated tempering\cite{
marinari1992, *lyubartsev1992}.
%with $i$ being the index of energy
%and that of temperature, respectively.
%
The estimated bias potentials at bin $i$, $V_i$,
are often improved
with the normalized histogram of the sampling, $H_i$, accumulated
during simulation.
%
Using $H_i$ as an estimate of $\pi_i$ in
Eq. \eqref{eq:pi_p_V}
allows us to correct $V_i$ as
%
\begin{equation}
  V^\mathrm{corr}_i
  =
  V_i
  +
  \ln \frac{ H_i }
           { p_i }.
  \label{eq:vcorr_equil}
\end{equation}
%
\note{This follows from
  $$
  \begin{aligned}
    H_i \approx \pi_i
    &\propto p^*_i \, e^{-V_i},
    \\
    p_i
    &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
  \end{aligned}
  $$
}
%However, for an unknown system,
%it is often difficult
%to guess a sufficiently accurate
%$V_i$ with which to start.
%%
%Thus,
%it can take many iterative rounds of simulations
%to achieve the desired flat distribution.
%%
%%Another difficulty of this approach is that
%The effectiveness of this approach
%depends on the precision of the histogram, $H_i$,
%which in turn depends on the simulation length
%that must be decided in advance.
However, this formula often requires a long
simulation period to accumulate a precise histogram, $H_i$,
and thus it disallows
continuous improvement of the bias potential
(hence that of the PMF)
as simulation lengthens.

The second type of methods,
including the WL algorithm and metadynamics,
address the above shortcomings
by continuously and incrementally updating the bias potential.
%For this reason, the second type of methods,
%referred to as the adaptive FDS methods,
%were invented.
%
These methods, refered to as the adaptive FDS methods,
will be discussed below.
%%
%%These methods incrementally update the bias potential
%%and can approximately sample the target distribution.
%%
%However,
%the on-the-fly updates break the microscopic reversibility
%of the underlying sampling mechanism.
%Further, failure to properly reduce the magnitude of
%the histogram-based increments
%would lead to roughness or errors in the final estimate of the PMF.
%%
%%Thus, for a conservative practitioner,
%%the E-FDS methods appear more appealing
%%once a sufficiently accurate bias potential is available.
%%
%To minimize this effect,
%many adaptive methods adopt a schedule to
%reduce the updating magnitude
%over the course of simulation\cite{
%marsili2006,
%belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
%liang2007,
%barducci2008}.
%%
%%Thus, in late stages of a long simulation,
%%the updating becomes negligible and
%%the adaptive simulation tends to
%%the corresponding equilibrium one.
%%
%%if the bias potential were converged to the exact value,
%%this adaptive FDS tends to the ideal equilibrium counterpart,
%%particular equilibrium simulation can be thought as
%%the optimal asymptotic limit of the adaptive counterpart,
%%
%For example, one can show that
%if the magnitude is controlled by
%the optimal inverse-time prescription\cite{
%belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
%liang2007,
%morozov2007, zhou2008,
%komura2012, *caparica2012, *caparica2014},
%the WL algorithm is equivalent to the ideal equilibrium counterpart
%in improving the precision of the bias potential,
%thus allowing for maximum efficiency
%(cf. Appendix \ref{sec:equilerr}).
%%the bias potential in the WL algorithm
%%will evolve similar to Eq. \eqref{eq:vcorr_equil}
%%with the precision of the histogram as simulation lengthens.

@@

%We will mainly focus on the adaptive methods below.
%
In the WL algorithm\cite{wang2001, *wang2001pre},
the bias potential, $V_i(t)$, is updated
at each sampling step $t$ as
%
\begin{equation}
  V_i(t+1)
  =
  V_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { p_{i(t)} }
  ,
\label{eq:wl_update}
\end{equation}
%
where $i(t)$ is the bin at step $t$,
$\delta_{i, \, i(t)}$ is the Kronecker delta function,
which is $1$ when $i = i(t)$ or $0$ otherwise,
and $\alpha(t)$ is the updating magnitude.
%
We refer to this scheme as the \emph{single-bin scheme},
for it applies only to the bias potential
at the current bin, $i(t)$.
%
In a more general \emph{multiple-bin scheme}, however,
several neighboring bins are updated:
%
\begin{equation}
  V_i(t+1)
  =
  V_i(t)
  +
  w_{i, \, i(t)}
  \frac{ \alpha(t) }
       { p_{i(t)} },
  \label{eq:mbin_update}
\end{equation}
%
where the elements, $w_{ij}$'s, are a set of coefficients satisfying
the restrictions to be discussed
in Sec. \ref{sec:updating-matrix}.
%
The Gaussian updating scheme
used in metadynamics
is a multiple-bin scheme.

@@

%For the single-bin scheme
%used in WL algorithm,
Later,
it was found that
a better way
of controlling the schedule, $\alpha(t)$,
in late stages
is to follow the formula
%
\begin{equation}
  \alpha(t) = \frac{1}{t},
  \label{eq:alpha_invt}
\end{equation}
%
where $t$,
referred to as the ``time'' below,
is the number of steps
from the beginning of the simulation.
%
This simple yet effective formula has attracted
several studies\cite{
belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
morozov2007, zhou2008,
komura2012, *caparica2012, *caparica2014}.
%
An interesting explanation\cite{
  marsili2006, barducci2008}
comes from the interpretation of
the bias potential %under the WL updating scheme, Eq. \eqref{eq:wl_update},
as a runtime average computed from all data points collected so far,
so that the addition of a new data point only
affects the average by its due weight, $1/t$,
which is the inverse of the current sample size, $t$
(cf. Appendix \ref{sec:equilerr}).
%
Here we will re-derive this formula
and suggest a slight modification
whereby the origin of $t$ is shifted
according to the initial error.
%
Interestingly, we may reach the same result
from optimally planning the simulation lengths in the stages
and the reduction factor, as shown in Appendix \ref{sec:wltoinvt}.

%As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
%several similar ways of
%reducing the updating magnitude
%were developed for umbrella sampling and metadynamics,
%particularly, the self-healing umbrella sampling
%and well-tempered metadynamics\cite{
%  marsili2006, barducci2008, dickson2011, dama2014}.
%%
Our aim here is to find the optimal schedule, $\alpha(t)$,
for a general updating scheme.

@@

First, we need to define the error in the bias potential.
%
This is done by deducting two contributions from the bias potential.
%
The first contribution comes from the target value,
Eq. \eqref{eq:Vi_target}.
%
For this purpose,
we introduce a shifted bias potential
%
\begin{equation}
  v_i(t)
  \equiv
  V_i(t)
  -
  \ln \frac { p^*_i }
            { p_i }
  ,
  \label{eq:v_def}
\end{equation}
%
which should tend to a constant of $i$
upon convergence.
Note that the updates to $v_i$'s are
equivalent to the updates to $V_i$'s,
since the shift
remains a constant during the course of simulation,
%%
%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
%becomes
%%
%\begin{equation}
%  \pi_i
%  =
%  \frac{                p_i \, e^{-v_i} }
%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
%  \propto
%  p_i \, e^{-v_i}.
%  \label{eq:pi_p_v}
%\end{equation}
%%
%According to Eq. \eqref{eq:pi_p_V},

@@

This is a functional of the curve $q(t)$.
%
We now fix the endpoint value, $q(T)$,
and vary the curve, $q(t)$ ($0 < t < T$).
%
This is equivalent to varying $u$
with fixed endpoints,
$u\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$u\bigl( q(T) \bigr) = 1$.
%
Note also that the first term on the right-hand side
of Eq. \eqref{eq:x2t_average} is fixed during
the process,
while the second term is the action of a free particle,
which must move at a constant velocity, $c$:
%
%We will further demand that
%%
%\begin{equation}
%  \lim_{t \to \infty} q(t) \to \infty.
%  \label{eq:qt_limit}
%\end{equation}
%%
%Then, for a large $t$,
%the second term on the right-hand side
%of Eq. \eqref{eq:xt_solution} can be neglected, and
%%
%\begin{align}
%\left\langle x^2(t) \right\rangle
%%&=
%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%%\notag
%%\\
%&=
%\int_0^t \int_0^t
%  \dot u\bigl( q(t') \bigr) \,
%  \dot u\bigl( q(t'') \bigr) \,
%  \kappa(t' - t'') \, dt'' \, dt'.
%\label{eq:x2t_average_asym}
%\end{align}
%
%By the Euler-Lagrange equation, we get
%$$
%\begin{aligned}
%0
%&=
%\frac{d}{dt} \int_0^T
%  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
%\\
%&= \int_0^T
%  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
%\end{aligned}
%$$
%where we have dropped the boundary terms
%by assuming Eq. \eqref{eq:kappa_limit}
%and a large $T$.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
%In order for this to hold for any $t$,
%we must have
%%
%$
%\ddot u\bigl( q(t') \bigr) = 0
%,
%$
%or
\begin{equation}
  \dot u\bigl( q(t') \bigr) = c
  .
\label{eq:dudt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:u_def}, we have
$$
e^{ q(t') - q(T) }
=
c \, (t' + t_0)
=
\frac{ t' + t_0 }
     { T  + t_0 }
,
$$
where $t_0$ is a constant,
and in the second step,
the constant $c$ is determined
from the boundary condition at $t' = T$.
%
Taking the logarithm,
and differentiating both sides
with respect to $t'$,
we get Eq. \eqref{eq:alpha_invt1}.
%
Further, from the case of $t' = 0$,
we get $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$.
%
Since the above optimization is performed
under a fixed $q(T)$,
we will complete the optimization
by choosing a proper value of $q(T)$,
or, equivalently, of $t_0$.
